TR Invariant T.I.

### 2017-04-19 Notes of various papers/TR invariant topological insulators T. Hughes. Dissertation/note.tex
[major] Calculate the Edge modes.

Questions:
\begin{itemize}
    \item Do we have a precise definition of \textit{topological phase
        transition}?
\end{itemize}

@@

Note, in choosing the test wavefunction, $u$ and $v$ could not be
seperated, because there is still interaction between the two
component in terms like $c^\dagger_{p_x,n}\sigma^x c_{p_x,n}$.
If we calculate $\tilde H_{LD}\ket{\psi_{p_x}} = E_{p_x}
\ket{\psi_{p_x}}$, we would get after careful calculation:

@@

Therefore, the eigenvalue equation could be turned into a matrix form:
\begin{align}
    H_{\text{disc}} \psi \equiv 
    \begin{pmatrix}
        B   & C &        &   & A \\
        A   & B & C      &   & \\
            & A & B      & C & \\
            &   & \cdots &   & \\
            &   & A      & B & C \\
        C   &   &        & A & B
    \end{pmatrix}
    \begin{pmatrix}
        \psi_{p_x,1} \\
        \psi_{p_x,2} \\
        \cdots \\
        \\
        \\
        \psi_{p_x,N} \\
    \end{pmatrix} =
    E_{p_x}
    \begin{pmatrix}
        \psi_{p_x,1} \\
        \psi_{p_x,2} \\
        \cdots \\
        \\
        \\
        \psi_{p_x,N} \\
    \end{pmatrix}
\end{align}

@@

    \subsection{Why I think the Lattice Model Hamiltonian is mildly wrong}
    \label{sec:Why-LatticeM-m-wrong}

    \subsection{Calculation Note I (Not related to the main discussion)}

@@

\bibliography{cite}{}
\bibliographystyle{alphaurl}

\subsection{Numerical Solution with open boundary condition in both
dimensions}
\label{sec:nsol-open-bc-inXY}

@@

\subsection{Edge States}
\label{sec:Edge States}
Inspired by lecture by Kane of Topological Insulators (BSS 2016), I
will analyse the continuum limit of this model at the  point
$(p_x=0,p_y=0)$, where the energy $E=0$. I also assume that we have placed
to materials adjacent to each other, on with $m<0$, and the other with
$m>0$. Therefore, at the interface, we could have $m$ as a function of
$x$ such that $m(0)=0$.

At this point $x=0$, $\sin(p_x)\to p_x$ , and is replaced by $-i\hbar
\partial _x$. The Schrodinger equation is H$\psi $=0 After the
calculation, the equation is (with $\hbar=1$):
\begin{equation}
    \left[i\partial_x \sigma^x + i\partial_y \sigma^y +
    m\sigma^z\right] \psi = 0
\end{equation}
However, this coupled PDE is hard to solve. Therefore, I restrict
their value on $x$, and solve the ODE:
\begin{align}
    i\partial_x \psi_2(x) + m\psi_1(x) &= 0 \\
    i\partial_x \psi_1(x) - m\psi_2(x) &= 0
\end{align}
If assuming $m(x)$ is in the form of $m(x)=x$, i.e. positive when
$x>0$, and negative when $x<0$, then the solution is:
\begin{align}
    \psi_1 (x,y) &= e^{\text{Int}(x)} C(y) \\
    \psi_2 (x,y) &= -i \psi_1(x,y)
\end{align}
where $\text{Int}(x)=\int_1^x -m(k)\dd{k}$. $\text{Int}(x)$ has the
property of goes to $-\infty$ as $x\to \pm\infty$.

@@

\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

@@

Note, in choosing the test wavefunction, $u$ and $v$ could not be
separated, because there is still interaction between the two
component in terms like $c^\dagger_{p_x,n}\sigma^x c_{p_x,n}$.
If we calculate $\tilde H_{LD}\ket{\psi_{p_x}} = E_{p_x}
\ket{\psi_{p_x}}$, we would get after careful calculation:

@@

Therefore, the eigenvalue equation could be turned into a matrix form:
\begin{align}
    \label{eq:disc-eigeneq}
    H_{\text{disc}} \psi \equiv 
    \begin{pmatrix}
        B   & C &        &   & A \\
        A   & B & C      &   & \\
            & A & B      & C & \\
            &   & \cdots &   & \\
            &   & A      & B & C \\
        C   &   &        & A & B
    \end{pmatrix}
    \begin{pmatrix}
        \psi_{p_x,1} \\
        \psi_{p_x,2} \\
        \cdots \\
        \\
        \\
        \psi_{p_x,N} \\
    \end{pmatrix} =
    E_{p_x}
    \begin{pmatrix}
        \psi_{p_x,1} \\
        \psi_{p_x,2} \\
        \cdots \\
        \\
        \\
        \psi_{p_x,N} \\
    \end{pmatrix}
\end{align}

@@

    \subsubsection{Calculation Note I (Not related to the main discussion)}

@@

\subsection{Why I Think the Lattice Model Hamiltonian is Mildly Wrong}
\label{sec:Why-LatticeM-m-wrong}

\subsection{Solution With Translational Invariance in x, and Open
Boundary in y (Infinite Stripe)}

This amounts to removing the $A$ in top right corner and $C$ in bottom
left corner of eq.\ref{eq:disc-eigeneq}. The calculation can be found
in file \texttt{Lattice Dirac Model (2+1)-d-4-Infinite Stripe.nb}.
However, the solution by Mathematica is analytically intractable and
not really illuminating. Also plotting all those eigenvectors becomes
time-consuming when $N$ is large. So I looked for similar things
inside the book by Bernevig \cite{Bernevig2013}, and found the
following model.

\subsection{Solution With Translational Invariance (Bloch States) in
both x and y (Infinite Plane)}

    \subsubsection{Numerical Solution}
    \label{sec:Numerical Solution}

@@

    \subsubsection{Edge States}
    \label{sec:Edge States}
    Inspired by lecture by Kane of Topological Insulators (BSS 2016), I
    will analyse the continuum limit of this model at the  point
    $(p_x=0,p_y=0)$, where the energy $E=0$. I also assume that we have placed
    to materials adjacent to each other in $x$ direction, on with
    $m<0$, and the other with $m>0$. Therefore, at the interface, we
    could have $m$ as a function of $x$ such that $m(0)=0$.

    At this point $x=0$, $\sin(p_x)\to p_x$ , and is replaced by $-i\hbar
    \partial _x$. So we recovered the continuum model:
    \begin{equation}
        H = -i\hbar\partial_x \sigma^x 
        -i\hbar\partial_y \sigma^y -m \sigma^z
    \end{equation}
    The Schrodinger equation is $H\psi =E\psi = 0$ for $E\approx 0$
    around this point. After some calculations, the equation is (with
    $\hbar=1$):
    \begin{equation}
        \left[i\partial_x \sigma^x + i\partial_y \sigma^y +
        m\sigma^z\right] \psi = 0
    \end{equation}
    However, this coupled PDE is hard to solve by Mathematica
    \footnote{See file \texttt{Lattice Dirac Model
    (2+1)-d-3-EdgeStates.nb} for my calculations.}. Therefore, I
    restrict their value on $x$, and solve the ODE:
    \begin{align}
        i\partial_x \psi_2(x) + m\psi_1(x) &= 0 \\
        i\partial_x \psi_1(x) - m\psi_2(x) &= 0
    \end{align}
    If assuming $m(x)$ is in the form of $m(x)=x$, i.e. positive when
    $x>0$, and negative when $x<0$, then the solution is:
    \begin{align}
        \psi_1 (x,y) &= e^{\text{Int}(x)} C(y) \\
        \psi_2 (x,y) &= -i \psi_1(x,y)
    \end{align}
    where $\text{Int}(x)=\int_1^x -m(k)\dd{k}$. $\text{Int}(x)$ has the
    property of goes to $-\infty$ as $x\to \pm\infty$.

@@

    Paused to think what I have got. I have in essential got a
    solution separated in $x,y$. Therefore, now I re-solve this edge
    mode, using technique of separation of variables 
    \footnote{Actually, I am just mimicking the calculation
    in section 8.8 of \cite{Bernevig2013}.}.

    Without loss of generality, assume $m(x)$ is like $x$, i.e. it
    goes to $\pm\infty$ as $x\to\pm\infty$. Then, separate the
    wavefunction into:
    \begin{equation}
        \psi(x,y) = \phi_1(x)\phi_2(y)
    \end{equation}
    Assuming we are solving a Schrodinger equation $H\psi=E\psi$, with
    $E$ being very small.  Now, inspired by previous calculation,
    write down directly $\phi_1(x)=e^{-\int_0^x m(x')\dd{x'}}$.  Then,
    the equation decoupled directly into:
    \begin{equation}
        \left(
            -i\sigma_y\partial_y + m(x)\left(i\sigma_x-\sigma_z\right)
        \right)\psi_2(y) = E\psi_2(y)
    \end{equation}
    Since $E$ is small, but $m(x)$ could be very large and is
    independent of $y$, we should make the term $(i\sigma_x-\sigma_z)$
    somehow disappear. This matrix is singular and has only eigenvalue
    $0$. Its eigenvector is $(i,1)^t$. So we assume
    $\phi_2(y)=\chi(x)(i,1)^t$.

    It turns out that $(i,1)^t$ is also the eigenvector of
    $\sigma_y$ with eigenvalue $-1$. So the equation becomes quite
    simple now:
    \begin{equation}
        i\partial_y\chi(y) = E\chi(y)
    \end{equation}
    The solution is $\chi(y)=e^{-iEy}$, a running wave in the $y$
    direction!
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\bibliography{../../Library}{}
\bibliographystyle{alphaurl}

\end{document}
